---
title: 'Lab 5 '
output:
  word_document:
    number_sections: true
date: "03-24-2025"
editor: 
  markdown: 
    wrap: 72
---

### Name: Matthew Trevino

In this lab, we will be using the `dplyr` package to explore student
evaluations of teaching data.

**You are expected to use functions from `dplyr` to do your data
manipulation!**

# Part 1: GitHub Workflow

Now that you have the Lab 5 repository cloned, you need to make sure you
can successfully push to GitHub. To do this you need to:

-   Open the `lab-5-student.qmd` file (in the lower right hand corner).
-   Change the `author` line at the top of the document (in the YAML) to
    your name.
-   Save your file either by clicking on the blue floppy disk or with a
    shortcut (command / control + s).
-   Click the "Git" tab in upper right pane
-   Check the "Staged" box for the `lab-5-student.qmd` file (the file
    you changed)
-   Click "Commit"
-   In the box that opens, type a message in "Commit message", such as
    "Added my name".
-   Click "Commit".
-   Click the green "Push" button to send your local changes to GitHub.

RStudio will display something like:

```         
>>> /usr/bin/git push origin HEAD:refs/heads/main
To https://github.com/atheobold/introduction-to-quarto-allison-theobold.git
   3a2171f..6d58539  HEAD -> main
```

Now you are ready to go! Remember, as you are going through the lab I
would strongly recommend rendering your HTML and committing your after
**every** question!

# Part 2: Some Words of Advice

Part of learning to program is learning from a variety of resources.
Thus, I expect you will use resources that you find on the internet.
There is, however, an important balance between copying someone else's
code and *using their code to learn*.

Therefore, if you use external resources, I want to know about it.

-   If you used Google, you are expected to "inform" me of any resources
    you used by **pasting the link to the resource in a code comment
    next to where you used that resource**.

-   If you used ChatGPT, you are expected to "inform" me of the
    assistance you received by (1) indicating somewhere in the problem
    that you used ChatGPT (e.g., below the question prompt or as a code
    comment), and (2) downloading and including the `.txt` file
    containing your **entire** conversation with ChatGPT.

Additionally, you are permitted and encouraged to work with your peers
as you complete lab assignments, but **you are expected to do your own
work**. Copying from each other is cheating, and letting people copy
from you is also cheating. Please don't do either of those things.

## Setting Up Your Code Chunks

-   The first chunk of this Quarto document should be used to *declare
    your libraries* (probably only `tidyverse` for now).
-   The second chunk of your Quarto document should be to *load in your
    data*.

## Save Regularly, Render Often

-   Be sure to **save** your work regularly.
-   Be sure to **render** your file every so often, to check for errors
    and make sure it looks nice.
    -   Make sure your Quarto document does not contain `View(dataset)`
        or `install.packages("package")`, both of these will prevent
        rendering.
    -   Check your Quarto document for occasions when you looked at the
        data by typing the name of the data frame. Leaving these in
        means the whole dataset will print out and this looks
        unprofessional. **Remove these!**
    -   If all else fails, you can set your execution options to
        `error: true`, which will allow the file to render even if
        errors are present.

# Part 3: Let's Start Working with the Data!

## The Data

The `teacher_evals` dataset contains student evaluations of reaching
(SET) collected from students at a University in Poland. There are SET
surveys from students in all fields and all levels of study offered by
the university.

The SET questionnaire that every student at this university completes is
as follows:

> Evaluation survey of the teaching staff of University of Poland.
> Please complete the following evaluation form, which aims to assess
> the lecturer’s performance. Only one answer should be indicated for
> each question. The answers are coded in the following way: 5 - I
> strongly agree; 4 - I agree; 3 - Neutral; 2 - I don’t agree; 1 - I
> strongly don’t agree.
>
> Question 1: I learned a lot during the course.
>
> Question 2: I think that the knowledge acquired during the course is
> very useful.
>
> Question 3: The professor used activities to make the class more
> engaging.
>
> Question 4: If it was possible, I would enroll for a course conducted
> by this lecturer again.
>
> Question 5: The classes started on time.
>
> Question 6: The lecturer always used time efficiently.
>
> Question 7: The lecturer delivered the class content in an
> understandable and efficient way.
>
> Question 8: The lecturer was available when we had doubts.
>
> Question 9. The lecturer treated all students equally regardless of
> their race, background and ethnicity.

These data are from the end of the winter semester of the 2020-2021
academic year. In the period of data collection, all university classes
were entirely online amid the COVID-19 pandemic. While expected learning
outcomes were not changed, the online mode of study could have affected
grading policies and could have implications for data.

**Average SET scores** were combined with many other variables,
including:

1.  **characteristics of the teacher** (degree, seniority, gender, SET
    scores in the past 6 semesters).
2.  **characteristics of the course** (time of day, day of the week,
    course type, course breadth, class duration, class size).
3.  **percentage of students providing SET feedback.**
4.  **course grades** (mean, standard deviation, percentage failed for
    the current course and previous 6 semesters).

This rich dataset allows us to **investigate many of the biases in
student evaluations of teaching** that have been reported in the
literature and to formulate new hypotheses.

Before tackling the problems below, study the description of each
variable included in the `teacher_evals_codebook.pdf`.

**1. Load the appropriate R packages for your analysis.**

```{r}
#| label: setup
#| echo: FALSE
# code chunk for loading packages
library(tidyverse)
library(dplyr)
library(naniar)
```

**2. Load in the `teacher_evals` data.**

```{r}
#| label: load-data
# code chunk for importing the data
teach_evals <- read_csv("teacher_evals.csv")
```

### Data Inspection + Summary

**3. Provide a brief overview (\~4 sentences) of the dataset.**

> The data has a rich assortment of various data on students and teachers based on specific classes. Some of this data includes variables like the teacher's academic degree, the amount of time they have worked, their gender, as well as stats about their population of students like the standard deviation of their grades or what percentage failed their class. 

```{r}
#| label: explore-data
# you may want to use code to answer this question
summary(teach_evals)
dim(teach_evals)
glimpse(teach_evals)
```

**4. What is the unit of observation (i.e. a single row in the dataset)
identified by?**

> Class_id, teacher_id, question_no. These three columns are used to create a specific variable per row. 

```{r}
#| label: row-identification
# you may want to use code to answer this question

```

**5. Use *one*`dplyr` pipeline to clean the data by:**

-   **renaming the `gender` variable `sex`**
-   **removing all courses with fewer than 10 respondents**
-   **changing data types in whichever way you see fit (e.g., is the
    instructor ID really a numeric data type?)**
-   **only keeping the columns we will use -- `course_id`, `teacher_id`,
    `question_no`, `no_participants`, `resp_share`, `SET_score_avg`,
    `percent_failed_cur`, `academic_degree`, `seniority`, and `sex`**

**Assign your cleaned data to a new variable named `teacher_evals_clean`
–- use these data going forward. Save the data as
`teacher_evals_clean.csv` in the `data-clean` folder.**

```{r}
#| label: data-cleaning
# renames "gender" to "sex" and filters no_participants for values > 10, keeps only variables we want
teacher_evals_clean <- teach_evals %>%
  rename(sex = gender) %>%
  filter(no_participants > 10) %>%
  mutate(teacher_id = as.character(teacher_id)) %>%
  select(course_id, teacher_id, question_no, no_participants, resp_share, 
         SET_score_avg, percent_failed_cur, academic_degree, seniority, sex)
write.csv(teacher_evals_clean, "data-clean/teacher_evals_clean.csv", row.names = FALSE)
```

**6. How many unique instructors and unique courses are present in the
cleaned dataset?**
There are 294 unique instructors and 921 unique courses. 

```{r}
#| label: unique-courses
# summarise with n_distinct to find the number of each
teacher_evals_clean |> 
summarise(unique_count = n_distinct(teacher_id)) 
teacher_evals_clean |> 
   summarise(unique_count = n_distinct(course_id))
```

**7. One teacher-course combination has some missing values, coded as
`NA`. Which instructor has these missing values? Which course? What
variable are the missing values in?**
> The missing values are in the percent_failed_cur column name for the PAB3SE004PA course taught by teacher_id 56347. 

```{r}
#| label: uncovering-missing-values
# display variables with n/a in them. manually saw values in data set
gg_miss_var(teacher_evals_clean)
```

**8. What are the demographics of the instructors in this study?
Investigate the variables `academic_degree`, `seniority`, and `sex` and
summarize your findings in \~3 complete sentences.**

> There are 4270 teacher's with doctorates, 1808 with masters, 467 with no degree, and 118 with professional degrees. 
The vast majority of the teachers (1639) have only two years of experience teaching. 3199 of the teachers are female and 3464 are male. 

```{r}
#| label: exploring-demographics-of-instructors
# code chunk for Q7
teacher_evals_clean %>%
  count(academic_degree)
teacher_evals_clean %>%
  count(seniority)
teacher_evals_clean %>%
  count(sex)

```

**9. Each course seems to have used a different subset of the nine
evaluation questions. How many teacher-course combinations asked all
nine questions?**

```{r}
#| label: teacher-course-asked-every-question
#I had someone help me with the filter and answer_all part because I didn't know about the %in% function
#filters question_no containing values 901 through 909
teacher_evals_clean %>%
  filter(question_no %in% 901:909) %>%
  #groups these values according to teacher id
  group_by(teacher_id) %>%  
  #summarise the values based on filter criteria listed above
  summarise(answered_all = all(question_no %in% 901:909)) %>% 
  #filters the questions to be true 
  filter(answered_all == TRUE) %>%  
  #shows how many teachers that meet the criteria are in the set
  summarise(count = n())
```

## Rate my Professor

**10. Which instructors had the highest and lowest average rating for
Question 1 (I learnt a lot during the course.) across all their
courses?**

> It appears that there are a lot of teachers with high average scores. I looked at the data set originally with the code before "highest <- result" to check this out, and a lot of the averages were 5 or 1. It was too many to list. 

```{r}
#| label: question-1-high-low
# code chunk for Q9
score_avg_question <- teacher_evals_clean %>%
  filter(question_no == 901) %>%
  group_by(teacher_id, course_id) %>%
  summarize(avg_score = mean(SET_score_avg, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(avg_score))
#"removes" the highest average score
highest <- score_avg_question %>% slice(1) 
#"removes" the lowest average score
lowest <- score_avg_question %>% slice(n())
#displays
highest
lowest
```

**11. Which instructors with one year of experience had the highest and
lowest average percentage of students failing in the current semester
across all their courses?**

> 106692 had the highest, while 102379, 103092, 106126, 86222, 98650, 98651 all shared the lowest average. 

```{r}
#| label: one-year-experience-failing-students
# code chunk for Q10
failing_percentage <- teacher_evals_clean %>%
  filter(seniority == 1) %>%
  group_by(teacher_id) %>%
  summarize(failing_percentage = mean(percent_failed_cur, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(failing_percentage))
```

**12. Which female instructors with either a doctorate or professional
degree had the highest and lowest average percent of students responding
to the evaluation across all their courses?**
> For those with doctorates, teacher 101508 had the highest response share of 52.1%. For those with professional degrees, teacher 80245 had the lowest response share of 7.7%. 

```{r}
#| label: female-instructor-student-response
# code chunk for Q11
average_doc_fem <- teacher_evals_clean %>%
  filter(sex == "female", academic_degree %in% c("dr", "prof")) %>%
  select(teacher_id, academic_degree, resp_share)
average_doc_fem
```
